---
title: "NHAHES_Data_Analysis"
author: "Tanmaya Tripathi"
date: "23 April 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 10
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: yes
  word_document: default
---

###__Libraries used__  
* library(sqldf)
* library(knitr)  
* library(ggplot2)  

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(sqldf)
library(knitr)
library(ggplot2)
options(scipen=999)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

###__Loading the Demographics Data__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ DEMO_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ DEMO_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ DEMO_F .csv", header = TRUE, sep = ",")
table_20072008 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2007-2008 _ DEMO_E .csv", header = TRUE, sep = ",")
table_20052006 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2005-2006 _ DEMO_D .csv", header = TRUE, sep = ",")
table_20032004 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2003-2004 _ DEMO_C .csv", header = TRUE, sep = ",")
table_20012002 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2001-2002 _ DEMO_B .csv", header = TRUE, sep = ",")
table_19992000 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 1999-2000 _ DEMO .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")]
table_20112012 <- table_20112012[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")]
table_20092010 <- table_20092010[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")]
table_20072008 <- table_20072008[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")]
table_20052006 <- table_20052006[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHINC","DMDEDUC2")]
table_20032004 <- table_20032004[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHINC","DMDEDUC2")]
table_20012002 <- table_20012002[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHINC","DMDEDUC2")]
table_19992000 <- table_19992000[,c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHINC","DMDEDUC2")]

colnames(table_19992000) <- c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")
colnames(table_20012002) <- c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")
colnames(table_20032004) <- c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")
colnames(table_20052006) <- c("RIDSTATR","RIAGENDR","SEQN","RIDAGEYR","RIDRETH1","DMDHHSIZ","INDHHIN2","DMDEDUC2")

demographics <- rbind(table_20132014
,table_20112012
,table_20092010
,table_20072008
,table_20052006
,table_20032004
,table_20012002
,table_19992000)
```

####__Columns important from demographics perspective__  
* RIDAGEYR - Age in years at screening  
* RIDRETH1 - Race/Hispanic origin w/ NH Asian  
* DMDHHSIZ - Total number of people in the Household  
* INDHHIN2 - Annual household income  
* DMDEDUC2 - Education level - Adults 20+  
  
__Filtering out the data based on the following conditions-__  
  
__1.__ We are keeping data where both interview and examination of the patient has happened.  
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, demographics$RIDSTATR == 2)
```
__2.__ For our analysis we will be working only with male patients.  
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, demographics$RIAGENDR == 1)
#demographics$RIAGENDR[demographics$RIAGENDR == 1]    <- "Male"
```
__3.__ The average urine flow rate for males age 14 to 45 is 21mL/sec. For our analysis we will be using the mentioned age group.  
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, demographics$RIDAGEYR > 13 & demographics$RIDAGEYR < 46)
```
__4.__ Renaming the race factors-  
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, !is.na(demographics$RIDRETH1))
#demographics$RIDRETH3[demographics$RIDRETH3 == 1]    <- "Mexican American"
#demographics$RIDRETH3[demographics$RIDRETH3 == 2]    <- "Other Hispanic"
#demographics$RIDRETH3[demographics$RIDRETH3 == 3]    <- "Non-Hispanic White"
#demographics$RIDRETH3[demographics$RIDRETH3 == 4]    <- "Non-Hispanic Black"
#demographics$RIDRETH3[demographics$RIDRETH3 == 6]    <- "Non-Hispanic Asian"
#demographics$RIDRETH3[demographics$RIDRETH3 == 7]    <- "Other Race - Including Multi-Racial"
```
__5.__ Factoring in Annual Household Income-  
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, !is.na(demographics$INDHHIN2))
demographics <- subset(demographics, demographics$INDHHIN2 != 77 & demographics$INDHHIN2 != 99)
#demographics$INDHHIN2[demographics$INDHHIN2 == 1]    <- "$0 to $4,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 2]    <- "$5,000 to $9,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 3]    <- "$10,000 to $14,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 4]    <- "$15,000 to $19,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 5]    <- "$20,000 to $24,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 6]    <- "$25,000 to $34,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 7]    <- "$35,000 to $44,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 8]    <- "$45,000 to $54,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 9]    <- "$55,000 to $64,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 10]    <- "$65,000 to $74,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 12]    <- "$20,000 and Over"
#demographics$INDHHIN2[demographics$INDHHIN2 == 13]    <- "Under $20,000"
#demographics$INDHHIN2[demographics$INDHHIN2 == 14]    <- "$75,000 to $99,999"
#demographics$INDHHIN2[demographics$INDHHIN2 == 15]    <- "$100,000 and Over"
```
__6.__ What is the highest education qualification for a patient-
```{r warning=FALSE, message=FALSE}
demographics <- subset(demographics, !is.na(demographics$DMDEDUC2))
demographics <- subset(demographics, demographics$DMDEDUC2 != 9 & demographics$DMDEDUC2 != 7)
#demographics$DMDEDUC2[demographics$DMDEDUC2 == 1]    <- "Less than 9th grade"
#demographics$DMDEDUC2[demographics$DMDEDUC2 == 2]    <- "9-11th grade (Includes 12th grade with no diploma)"
#demographics$DMDEDUC2[demographics$DMDEDUC2 == 3]    <- "High school graduate/GED or equivalent"
#demographics$DMDEDUC2[demographics$DMDEDUC2 == 4]    <- "Some college or AA degree"
#demographics$DMDEDUC2[demographics$DMDEDUC2 == 5]    <- "College graduate or above"
```

###__Loading the Urine Lab Data__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ UCFLOW_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ UCFLOW_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ UCFLOW_F .csv", header = TRUE, sep = ",")

urine_lab <- rbind(table_20132014
,table_20112012
,table_20092010)
```

####__Columns important from urine flow data perspective__  
* URDFLOW1 - Urine #1 Flow Rate  

__Filtering out the data based on the following conditions-__  
  
__1.__ We are keeping data where urine flow rate data is present.  
```{r warning=FALSE, message=FALSE}
urine_lab <- subset(urine_lab, !is.na(urine_lab$URDFLOW1))
urine_lab$URDFLOW1_STATE <- ifelse(urine_lab$URDFLOW1 > 0.35, "Nor", "Abnor")
```

###__Loading the Alcohol Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ ALQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ ALQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ ALQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","ALQ101")]
table_20112012 <- table_20112012[,c("SEQN","ALQ101")]
table_20092010 <- table_20092010[,c("SEQN","ALQ101")]

alcohol_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

alcohol_questionairre_data$ALQ101 <- ifelse(is.na(alcohol_questionairre_data$ALQ101), 2, alcohol_questionairre_data$ALQ101)
alcohol_questionairre_data <- subset(alcohol_questionairre_data, alcohol_questionairre_data$ALQ101 != 9)
```

We have imputed the null values here with the patient saying that they do not have 12 drinks or more in a year. Therefore, for our understanding null value signifies a no (2) while yes is 1.  

###__Loading the Blood Pressure and Cholestrol Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ BPQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ BPQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ BPQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","BPQ020","BPQ080")]
table_20112012 <- table_20112012[,c("SEQN","BPQ020","BPQ080")]
table_20092010 <- table_20092010[,c("SEQN","BPQ020","BPQ080")]

bp_chol_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

bp_chol_questionairre_data <- subset(bp_chol_questionairre_data, bp_chol_questionairre_data$BPQ080 != 9)
bp_chol_questionairre_data <- subset(bp_chol_questionairre_data, bp_chol_questionairre_data$BPQ020 != 9)
```

###__Loading the Dermatology Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ DEQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ DEQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ DEQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","DED031","DEQ034A","DEQ034D")]
table_20112012 <- table_20112012[,c("SEQN","DED031","DEQ034A","DEQ034D")]
table_20092010 <- table_20092010[,c("SEQN","DED031","DEQ034A","DEQ034D")]

dermatology_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

dermatology_questionairre_data <- subset(dermatology_questionairre_data, dermatology_questionairre_data$DED031 !=77 & dermatology_questionairre_data$DED031 != 99)
dermatology_questionairre_data <- subset(dermatology_questionairre_data, dermatology_questionairre_data$DEQ034A !=77 & dermatology_questionairre_data$DEQ034A != 99)
dermatology_questionairre_data <- subset(dermatology_questionairre_data, dermatology_questionairre_data$DEQ034D !=77 & dermatology_questionairre_data$DEQ034D != 99)

dermatology_questionairre_data <- subset(dermatology_questionairre_data, !is.na(dermatology_questionairre_data$DED031))
dermatology_questionairre_data <- subset(dermatology_questionairre_data, !is.na(dermatology_questionairre_data$DEQ034A))
dermatology_questionairre_data <- subset(dermatology_questionairre_data, !is.na(dermatology_questionairre_data$DEQ034D))
```

###__Loading the Diabetes Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ DIQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ DIQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ DIQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","DIQ010","DIQ050")]
table_20112012 <- table_20112012[,c("SEQN","DIQ010","DIQ050")]
table_20092010 <- table_20092010[,c("SEQN","DIQ010","DIQ050")]

diabetes_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

diabetes_questionairre_data <- subset(diabetes_questionairre_data, diabetes_questionairre_data$DIQ010 <= 3)
diabetes_questionairre_data <- subset(diabetes_questionairre_data, diabetes_questionairre_data$DIQ050 <= 2)

diabetes_questionairre_data <- subset(diabetes_questionairre_data, !is.na(diabetes_questionairre_data$DIQ010))
diabetes_questionairre_data <- subset(diabetes_questionairre_data, !is.na(diabetes_questionairre_data$DIQ050))
```

###__Loading the Diet Behavior Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ DBQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ DBQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ DBQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","DBQ197","DBQ229","DBQ235A","DBQ235B","DBQ235C","DBD905","DBD910")]
table_20112012 <- table_20112012[,c("SEQN","DBQ197","DBQ229","DBQ235A","DBQ235B","DBQ235C","DBD905","DBD910")]
table_20092010 <- table_20092010[,c("SEQN","DBQ197","DBQ229","DBQ235A","DBQ235B","DBQ235C","DBD905","DBD910")]

diet_beh_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBQ197 <= 4)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBQ229 <= 3)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBQ235A <= 3)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBQ235B <= 3)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBQ235C <= 3)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBD905 <= 200)
diet_beh_questionairre_data <- subset(diet_beh_questionairre_data, diet_beh_questionairre_data$DBD910 <= 200)
```

###__Loading the Drug Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ DUQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ DUQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ DUQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","DUQ217","DUQ219")]
table_20112012 <- table_20112012[,c("SEQN","DUQ217","DUQ219")]
table_20092010 <- table_20092010[,c("SEQN","DUQ217","DUQ219")]

drug_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

drug_questionairre_data$DUQ217 <- ifelse(is.na(drug_questionairre_data$DUQ217), 0, drug_questionairre_data$DUQ217)
drug_questionairre_data$DUQ219 <- ifelse(is.na(drug_questionairre_data$DUQ219), 0, drug_questionairre_data$DUQ219)

drug_questionairre_data <- subset(drug_questionairre_data, drug_questionairre_data$DUQ217 <= 5)
drug_questionairre_data <- subset(drug_questionairre_data, drug_questionairre_data$DUQ219 <= 5)
```

###__Loading the Health Insurance Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ HIQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ HIQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ HIQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","HIQ011")]
table_20112012 <- table_20112012[,c("SEQN","HIQ011")]
table_20092010 <- table_20092010[,c("SEQN","HIQ011")]

health_insurance_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

health_insurance_questionairre_data <- subset(health_insurance_questionairre_data, health_insurance_questionairre_data$HIQ011 <= 5)
```

###__Loading the Hospital Utilization Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ HUQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ HUQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ HUQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","HUQ010","HUQ030")]
table_20112012 <- table_20112012[,c("SEQN","HUQ010","HUQ030")]
table_20092010 <- table_20092010[,c("SEQN","HUQ010","HUQ030")]

hospital_util_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

hospital_util_questionairre_data <- subset(hospital_util_questionairre_data, hospital_util_questionairre_data$HUQ010 <= 5)
hospital_util_questionairre_data <- subset(hospital_util_questionairre_data, hospital_util_questionairre_data$HUQ030 <= 3)
```

###__Loading the Housing Char Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ HOQ_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ HOQ_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ HOQ_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","HOQ065")]
table_20112012 <- table_20112012[,c("SEQN","HOQ065")]
table_20092010 <- table_20092010[,c("SEQN","HOQ065")]

housing_char_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

housing_char_questionairre_data <- subset(housing_char_questionairre_data, housing_char_questionairre_data$HOQ065 <= 3)
```

###__Loading the Kidney Conditions Questionairre Interview Data Set__
```{r warning=FALSE, message=FALSE}
table_20132014 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2013-2014 _ KIQ_U_H .csv", header = TRUE, sep = ",")
table_20112012 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2011-2012 _ KIQ_U_G .csv", header = TRUE, sep = ",")
table_20092010 <- read.csv(file = "M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/ 2009-2010 _ KIQ_U_F .csv", header = TRUE, sep = ",")

table_20132014 <- table_20132014[,c("SEQN","KIQ022","KIQ026")]
table_20112012 <- table_20112012[,c("SEQN","KIQ022","KIQ026")]
table_20092010 <- table_20092010[,c("SEQN","KIQ022","KIQ026")]

kidney_cond_questionairre_data <- rbind(table_20132014
,table_20112012
,table_20092010)

kidney_cond_questionairre_data <- subset(kidney_cond_questionairre_data, kidney_cond_questionairre_data$KIQ022 <= 2)
kidney_cond_questionairre_data <- subset(kidney_cond_questionairre_data, kidney_cond_questionairre_data$KIQ026 <= 2)
```

###__Joining Data__  
The below sql query is used to join the data.  
```{r warning=FALSE, message=FALSE}
complete_data <- sqldf("SELECT de.SEQN,
                               de.RIDAGEYR,
                               de.RIDRETH1,
                               de.DMDHHSIZ,
                               de.INDHHIN2,
                               de.DMDEDUC2,
                               aqd.ALQ101,
                               bpqd.BPQ020,
                               bpqd.BPQ080,
                               dqd.DED031,
                               dqd.DEQ034A,
                               dqd.DEQ034D,
                               diqd.DIQ010,
                               diqd.DIQ050,
                               dbqd.DBQ197,
                               dbqd.DBQ229,
                               dbqd.DBQ235A,
                               dbqd.DBQ235B,
                               dbqd.DBQ235C,
                               dbqd.DBD905,
                               dbqd.DBD910,
                               duqd.DUQ217,
                               duqd.DUQ219,
                               hiqd.HIQ011,
                               huqd.HUQ010,
                               huqd.HUQ030,
                               hcqd.HOQ065,
                               kcqd.KIQ022,
                               kcqd.KIQ026,
                               la.URDFLOW1,
                               la.URDFLOW1_STATE
                        FROM   demographics de LEFT OUTER JOIN urine_lab la ON de.SEQN = la.SEQN
                                               LEFT OUTER JOIN alcohol_questionairre_data aqd ON de.SEQN = aqd.SEQN
                                               LEFT OUTER JOIN bp_chol_questionairre_data bpqd ON de.SEQN = bpqd.SEQN
                                               LEFT OUTER JOIN dermatology_questionairre_data dqd ON de.SEQN = dqd.SEQN
                                               LEFT OUTER JOIN diabetes_questionairre_data diqd ON de.SEQN = diqd.SEQN
                                               LEFT OUTER JOIN diet_beh_questionairre_data dbqd ON de.SEQN = dbqd.SEQN
                                               LEFT OUTER JOIN drug_questionairre_data duqd ON de.SEQN = duqd.SEQN
                                               LEFT OUTER JOIN health_insurance_questionairre_data hiqd ON de.SEQN = hiqd.SEQN
                                               LEFT OUTER JOIN hospital_util_questionairre_data huqd ON de.SEQN = huqd.SEQN
                                               LEFT OUTER JOIN housing_char_questionairre_data hcqd ON de.SEQN = hcqd.SEQN
                                               LEFT OUTER JOIN kidney_cond_questionairre_data kcqd ON de.SEQN = kcqd.SEQN")

complete_data <- subset(complete_data, !is.na(complete_data$URDFLOW1))

for(i in 1:ncol(complete_data)){
  if(nrow(complete_data[is.na(complete_data[,i]),])>0)
    print(colnames(complete_data[i]))
}

complete_data <- subset(complete_data, !is.na(complete_data$BPQ020))
complete_data <- subset(complete_data, !is.na(complete_data$DED031))
complete_data <- subset(complete_data, !is.na(complete_data$DIQ010))
complete_data <- subset(complete_data, !is.na(complete_data$DBQ197))
complete_data <- subset(complete_data, !is.na(complete_data$DUQ217))
complete_data <- subset(complete_data, !is.na(complete_data$HIQ011))
complete_data <- subset(complete_data, !is.na(complete_data$HUQ010))
complete_data <- subset(complete_data, !is.na(complete_data$HOQ065))
table(complete_data$URDFLOW1_STATE)

write.csv(complete_data, paste("M:/Documents/ML_Healthcare_Final_Project/NHANES_Dataset/complete_data.csv"))
```











